Fast Sorting on Modern Computers

### 2005-07-31 docs/tech_report/conclusions.tex
some spell checks, removed unneccessary files, rename elementary

\subsection{Heapsorts}

@@

\subsection{Mergesorts}

@@

\subsection{Quicksorts}

Our results show base quicksort being the fastest of the quicksorts. We note our
lowest instruction count occurs with the median-of-7 partitioning. We also see
that using a binary search slows multi-quicksort down compared to a sequential
search, but that its extra instructions make it slower than tuned quicksort.
Despite being slower in our tests, pur simulations show tuned quicksort to have
fewer cache misses, instructions and branch mispredictions than base quicksort.

\subsection{Radixsorts}

@@

\subsection{Shellsorts}

@@

Insertion sort has exaclty one miss per key. Using this result we were able to
make several other sorts faster by replacing their simple inline sorts with
insertion sorts.

Selection sort also has a very low number of branch mispredictions, the results
of which match up well to Knuth's calculation. We found aswell that bubblesort
and shakersort have incredibly bad branch prediction performance, bring
$O(N^2)$.

@@

Over the course of this research. The vast majority of our results have
correlated exactly with LaMarca's. Many of our graphs are exactly the same
shape to LaMarca's, can be calculated exactly or alomst exactly using his
methods, or can be seen to be equivalent when scaled by an appropriate ratio
(due to different test parameters).

@@

LaMarca's initial improvements to base mergesort were due to reduce the
instruction count. We noted a reduction with each of the steps he prescribed. We
note too that each of his improvement reduce level 2 cache misses, and the shape
of our level 2 cache miss graphs exactly match his. We also note similar
characterists: LaMarca mentions a wobble in the instruction count graphs, due to
having to copy the final merge back; we note this same wobble.

However, we disagree with his results in several key places. Firstly, we note
that an inlined sorting method does not perform as fast as insertion sort. We
find it difficult to perform his improvements on either algorithm N or algorithm
S, and write our own mergesort instead. We expected that mergesort would have
half the instruction count of heapsortl we found instead they were very similar.

LaMarca did not include details of his tagging system in multi-mergesort. As a
result, we used a search, which greatly increased the instruction count. As a
result, mutli-mergesort did not perform as well as the mergesorts which had only
been tiled.

@@

Where LaMarca found that using Sedgewicks insertion sort optimisation decreased
instruction count, we found the opposite. We also found that despite
improvements due to cache performance and instruction count, our base quicksort
performed faster than our memory-tuned version. Finally, we found that although
changing multi-mergesort's sequential search to a binary search decreased the
instruction count, it greatly increased the number of branch mispredictions
incurred, and slowed the sort down accordingly.

>>>>

\subsection{Heapsort}

@@

\subsection{Mergesort}

@@

\subsection{Quicksort}

Our results show base quicksort being the fastest of the quicksorts. We note our
lowest instruction count occurs with the median-of-7 partitioning. We also see
that using a binary search slows multi-quicksort down compared to a sequential
search, but that its extra instructions make it slower than memory-tuned
quicksort.  Despite being slower in our tests, our simulations show tuned
quicksort to have fewer cache misses, instructions and branch mispredictions
than base quicksort.

\subsection{Radixsort}

@@

\subsection{Shellsort}

@@

Insertion sort has exactly one miss per key. Using this result we were able to
make several other sorts faster by replacing their simple inline sorts with
insertion sorts.

Selection sort also has a very low number of branch mispredictions, the results
of which match up well to Knuth's calculation. We found as well that bubblesort
and shakersort have incredibly bad branch prediction performance, bring
$O(N^2)$.

@@

Over the course of this research. The vast majority of our results have
correlated exactly with LaMarca's. Many of our graphs are exactly the same
shape to LaMarca's, can be calculated exactly or almost exactly using his
methods, or can be seen to be equivalent when scaled by an appropriate ratio
(due to different test parameters).

@@

LaMarca's initial improvements to base mergesort were due to reduce the
instruction count. We noted a reduction with each of the steps he prescribed. We
note too that each of his improvement reduce level 2 cache misses, and the shape
of our level 2 cache miss graphs exactly match his. We also note similar
characteristics: LaMarca mentions a wobble in the instruction count graphs, due to
having to copy the final merge back; we note this same wobble.

However, we disagree with his results in several key places. Firstly, we note
that an in-lined sorting method does not perform as fast as insertion sort. We
find it difficult to perform his improvements on either algorithm N or algorithm
S, and write our own mergesort instead. We expected that mergesort would have
half the instruction count of heapsortl we found instead they were very similar.

LaMarca did not include details of his tagging system in multi-mergesort. As a
result, we used a search, which greatly increased the instruction count. As a
result, multi-mergesort did not perform as well as the mergesorts which had only
been tiled.

@@

Where LaMarca found that using Sedgewick's insertion sort optimisation decreased
instruction count, we found the opposite. We also found that despite
improvements due to cache performance and instruction count, our base quicksort
performed faster than our memory-tuned version. Finally, we found that although
changing multi-mergesort's sequential search to a binary search decreased the
instruction count, it greatly increased the number of branch mispredictions
incurred, and slowed the sort down accordingly.
